---
title: "Regression Models Course Project"
author: "EricRybicki"
date: "24 Jan 2015"
output: html_document
---

A brief investigation into miles per gallon efficiency of 32 cars by transmission type.

### Description
```{r}
mtcars <- mtcars
?mtcars
```
The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

### Format

A data frame with 32 observations on 11 variables.
`[, 1]  mpg     Miles/(US) gallon`

`[, 2]  cyl	    Number of cylinders`

`[, 3]	disp	Displacement (cu.in.)`

`[, 4]	hp	    Gross horsepower`

`[, 5]	drat	Rear axle ratio`

`[, 6]	wt	    Weight (lb/1000)`

`[, 7]	qsec	1/4 mile time`

`[, 8]	vs	    V/S`

`[, 9]	am	    Transmission (0 = automatic, 1 = manual)`

`[,10]	gear	Number of forward gears`

`[,11]	carb	Number of carburetors`



## Is an automatic or manual transmission better for MPG?


As we're focusing on the transmission type, we first want to re-label the `am` values to remove any ambiguity.
```{r}
mtcars$am[mtcars$am == 0] <- "Automatic"
mtcars$am[mtcars$am == 1] <- "Manual"
```

Next we take a quick look at the data to see what the shape of the data might tell us. 

We can clearly see from the plot in `fig.1` the manual transmission has a higher value spread of MPG compared to the automatic transmission. 


## Quantify the MPG difference between automatic and manual transmissions.
As there are a number of variables in the data set which might contribute to the MPG variances we can begin simply with a sole predictive transmission factor. `Fig.3`

We see that transmission alone cannot fully account for MPG. Intuitively we might infer that `weight` might play a role but and we can consider other significant factors. `Fig.4`
```{r, echo=FALSE}
am.model <- lm(mpg ~ factor(am), data = mtcars)
```

```{r, echo=FALSE}
qsec.model <- lm(mpg ~ wt + qsec + factor(am), data = mtcars)
```
From `Fig.5` we see an r-squared value of 0.8499 which shows us that this model is responsible for 85% of the variance in `MPG`
```{r}
anova(am.model, qsec.model)
```
With a p-value of `1.5e-09` we can deduce that this model significantly differs from our base model and can be used as a reasonable approximation of MPG influence. We can thus conclude from `summary(hp.model)` that an expected `2.936 MPG` can be attributed to a manual transmission over an automatic transmission. 

Finally we check the residuals with the plot in `Fig.3` for any signs of non-normality and find no discernible patterns which might not satisfy our assumptions.

## Appendix

##### Fig.1
```{r, message=FALSE}
require(gridExtra)
library(ggplot2)
bg <- ggplot(mtcars, aes(factor(am), mpg)) + geom_boxplot(colour="seagreen", size = 1.25) + 
            geom_jitter() + theme_bw() + xlab("Transmission")
vg <- ggplot(mtcars, aes(factor(am), mpg)) + geom_violin(colour="firebrick", size = 1.25) +
            geom_jitter() + theme_bw() + xlab("Transmission")

grid.arrange(bg, vg, ncol=2)
```

##### Fig.2
```{r}
am.model <- lm(mpg ~ factor(am), data = mtcars)
summary(am.model)
```

##### Fig.3
```{r}
hp.model <- lm(mpg ~ wt + hp + factor(am), data = mtcars)
summary(hp.model)
```

##### Fig.4
```{r}
qsec.model <- lm(mpg ~ wt + qsec + factor(am), data = mtcars)
summary(qsec.model)
```

##### Fig.5
```{r}
par(mfrow = c(2,2))
plot(qsec.model)
```







